### 2016-03-15 paper.tex
Add Wosnik edits to results section

Mean rotor power coefficients for multiple Reynolds numbers are plotted versus
tip speed ratio in Fig.~\ref{fig:cp-curves}. In general, $C_P$ increases with
$Re$, along with a reduction in the optimal tip speed ratio, due to the tendency
of foils to stall at higher angles of attack at higher $Re$, and the higher
angle of attack ranges seen by the blades at lower $\lambda$. These effects
diminish with increasing $Re$, which is expected as the blade boundary layers
transition to turbulence closer to the leading edge \cite{Lissaman1983,
    McMasters1980, Bachant2016-RVAT-Re-dep}, which helps flow remain attached longer
as it moves against the adverse pressure gradient on the suction side of the
foil. Note how at lower tip speed ratios and Reynolds numbers, $C_P$ becomes
significantly negative, which was possible thanks to the speed control of the
experimental setup's servo motor applying negative torque to the rotor.

@@

Mean rotor drag coefficients are plotted versus tip speed ratio in
Fig.~\ref{fig:cd-curves}. These $C_D$ curves show fewer differences compared to
the power coefficient curves, which is most noticeable at low $\lambda$ and low
$Re$. The relative similarity could be attributed to the effects of stall, where
the lift-to-drag ratio on the blades may drop, decreasing rotor torque, though
the total resultant force due to high blade drag retains a similar component in
the streamwise direction.

@@

The behavior of the power and drag coefficients at $\lambda=3.1$ are plotted in
Fig.~\ref{fig:perf-re-dep}, where we can see a Reynolds number independent value
of $C_D$ was observed. The power coefficient of the turbine increases
dramatically below $Re_D = 1 \times 10^6$ or $Re_c \equiv \lambda U_\infty c /
\nu \approx 2 \times 10^5$, beyond which there appears to be a small, linear,
positive trend. At the lowest Reynolds number, mean power coefficient even drops
below zero, which is consistent with the low performance of the 1:15 scale RM2
physical model study~\cite{Hill2014}. However, the blades of the 1:15 scale RM2
were mounted at approximately 58\% of the chord from the leading edge, versus
50\% for the present model, which helps explain the 1:15 scale model's positive
power output at $\lambda=2.2$ and $Re_D \sim 10^5$, since moving the blade mount
point further back is equivalent to a ``toe-out'' preset pitch
condition~\cite{Fiedler2009}.

The tendency of $C_P$ to continue increasing slightly could be
an effect of flow curvature---caused by the finite $c/R$---which imparts a
``virtual camber'' \cite{Migliore1980}, or can be thought of as producing a
``lead'' in angle of attack since cambered airfoils have non-zero lift at zero
angle of attack \cite{Goude2012}. The weaker $Re$-independence of the RM2
compared to that of the higher $c/R$ UNH-RVAT could be due to this virtual
camber effect, since camber has been shown to cause earlier $Re$-convergence of
the CFT's geometric torque coefficient when calculated from static foil
coefficients given by a viscous panel method, whereas characteristics like
lift-to-drag ratio do not converge as strongly \cite{Bachant2016-RVAT-Re-dep}.

@@

Performance curves for the rotor with both NACA 0021 and cylindrical struts are
shown in Fig.~\ref{fig:perf-covers}. As expected, the high drag cylindrical
struts reduce performance dramatically, which is more pronounced as $\lambda$
increases. This is in accordance with the measurements of Rawlings
\cite{Rawlings2008}, though the strut losses in this case are larger due to the
very high drag profile.

@@

Measurements for the power coefficient contributions of the strut drag losses
are presented in Fig.~\ref{fig:no-blades} for NACA 0021 and cylindrical
struts---both in towed and stationary conditions. These are computed in the same
fashion as the curves in Fig.~\ref{fig:cp-curves}, but with the blades removed.
We see that strut drag losses increase with tip speed ratio to the power 2--3,
which makes streamlined struts much more important for low solidity turbines,
given the inverse correlation between solidity and $\lambda_0$
\cite{Templin1974}.

@@

Strut drag losses did not change much for the streamlined NACA 0021 struts in
the towed versus stationary configuration, which helps explain why overall rotor
drag coefficients remained relatively constant even at low Reynolds number,
where $C_P$ was dramatically reduced, or even negative. For the cylindrical
struts, losses increased significantly when towed and operating in the mid range
of tip speed ratios. Though a real turbine would never use such high drag
struts, with respect to numerical modeling, these measurements will allow
support struts to be simulated without blades, to isolate and evaluate the
ability to predict these losses independent of the blade loading.

@@

The mean velocity field at the chosen optimal tip speed ratio $\lambda_0=3.1$, 1
m downstream ($x/D=0.93$) is plotted in Fig.~\ref{fig:meancontquiv}. The mean
streamwise velocity deficit is markedly more symmetric than that of the
UNH-RVAT, shown in Fig.~\ref{fig:rvat-meancontquiv}, with the RM2 inducing lower
acceleration around the turbine due to the lower rotor drag coefficient and a
slightly lower blockage ratio \cite{Bachant2015-JoT}. Tip vortex shedding is
relatively weaker, which is likely an effect of the RM2's smaller blade chord
length and tapered blades.

@@

Analysis of the near-wake of the higher solidity UNH-RVAT turbine revealed that
vertical advection was the largest contributor to streamwise recovery
\cite{Bachant2015-JoT}, a trait which is considered an advantage of cross-flow
over axial-flow rotors in arrays \cite{Kinzel2012}. A similar analysis was
undertaken for the RM2, starting with the transport equation for mean kinetic
energy, rearranged to isolate the streamwise partial derivative (indices follow
the Einsteinian convention):
\begin{equation}
    \begin{split}
        \frac{\p K}{\p x}
        =
        \frac{1}{U}
        \bigg{[}
        & - \underbrace{V \frac{\p K}{\p y}}_{y\text{-adv.}}
        - \underbrace{W \frac{\p K}{\p z}}_{z\text{-adv.}}
        % Pressure work:
        - \frac{1}{\rho}\frac{\p}{\p x_j} P U_i \delta_{ij}
        % Work by viscous forces
        + \frac{\p}{\p x_j} 2 \nu U_i S_{ij}
        % Turbulent transport of K
        - \underbrace{
            \frac{1}{2}\frac{\p}{\p x_j} \overline{u_i' u_j'} U_i
        }_{\text{Turb. trans.}} \\
        % Production of k
        & + 
        \underbrace{
            \overline{u_i' u_j'} \frac{\p U_i}{\p x_j}
        }_{k\text{-prod.}}
        % Mean dissipation
        - 
        \underbrace{
            2 \nu S_{ij}S_{ij}
        }_{\text{Mean diss.}}
        \bigg{]}.
    \end{split}
    \label{eq:K-full}
\end{equation}

@@

Derivatives were computed with a second order central difference scheme for
interior points, and a second order inward-facing scheme for the edges,
identical to \cite{Bachant2015-JoT}. Weighted averages for these computation are
shown in Fig.~\ref{fig:Ktransport}. Due to the weaker blade vortex shedding,
transport due to vertical advection at this point in the wake was approximately
3 times lower than the higher solidity UNH-RVAT. Note that direct comparison is
somewhat invalid, since the total measurement plane area was about 5\% lower for
this experiment. However, the differences observed are beyond this discrepancy.
We also see relatively lower levels of cross-stream turbulent transport due to
the lack of blade stall vortex shedding. These results may have interesting
implications regarding the application of turbines with lower power coefficient
to possibly improve overall array performance through enhanced transport of
kinetic energy from the free stream, though evaluating these trade-offs will
require a detailed analysis of the downstream evolution and turbine--wake
interaction.

>>>>

Mean rotor power coefficients for multiple Reynolds numbers are plotted versus
tip speed ratio in Fig.~\ref{fig:cp-curves}. In general, $C_P$ increases with
$Re$, along with a reduction in the optimal tip speed ratio, due to the tendency
of foils to stall at higher angles of attack at higher $Re$, and the higher
angle of attack ranges seen by the blades at lower $\lambda$. These effects
diminish with increasing $Re$, which is expected as the blade boundary layers
transition to turbulence closer to the leading edge \cite{Lissaman1983,
    McMasters1980, Bachant2016-RVAT-Re-dep}, which helps flow remain attached longer
as it moves against the adverse pressure gradient on the suction side of the
foil. For the experiment reported here, a maximum power coefficient $C_P=0.37$
was reached at $Re_D=1.3 \times 10^6$. Note how at lower tip speed ratios and
Reynolds numbers, $C_P$ becomes significantly negative, which was possible
thanks to the speed control of the experimental setup's servo motor applying
negative torque to the rotor.

@@

Mean rotor drag coefficients are plotted versus tip speed ratio in
Fig.~\ref{fig:cd-curves}. These $C_D$ curves show little difference compared to
the power coefficient curves, which is most noticeable at low $\lambda$ and low
$Re$. The relative similarity could be attributed to the effects of stall, where
the lift-to-drag ratio on the blades may drop, decreasing rotor torque, though
the total resultant force due to high blade drag retains a similar component in
the streamwise direction. When operating at maximum power coefficient,
$\lambda_0=3.1$, and $Re_D=1.3 \times 10^6$, a rotor drag coefficient $C_D=0.84$
was measured.

@@

The effects of Reynolds number on the power and drag coefficients at
$\lambda=3.1$ are shown in Fig.~\ref{fig:perf-re-dep}. The rotor drag
coefficient $C_D$ became more or less Reynolds number independent for $Re_D \ge
0.6 \times 10^6$. The power coefficient of the turbine increased dramatically
below $Re_D = 1 \times 10^6$ or $Re_c \equiv \lambda U_\infty c / \nu \approx 2
\times 10^5$, beyond which there appears to be a small, linear, positive trend.
At the lowest Reynolds number, mean power coefficient even dropped below zero,
which is consistent with the low performance of the 1:15 scale RM2 physical
model study~\cite{Hill2014}. However, the blades of the 1:15 scale RM2 were
mounted at approximately 58\% of the chord from the leading edge, versus 50\%
for the present model, which helps explain the 1:15 scale model's positive power
output at $\lambda=2.2$ and $Re_D \sim 10^5$, since moving the blade mount point
further back is equivalent to a ``toe-out'' preset pitch
condition~\cite{Fiedler2009}.

The tendency of $C_P$ to continue increasing slightly could be an effect of flow
curvature---caused by the finite $c/R$---which imparts a ``virtual camber''
\cite{Migliore1980}, or can be thought of as producing a ``lead'' in angle of
attack since cambered airfoils have non-zero lift at zero angle of attack
\cite{Goude2012}. The residual weak $Re$-dependence of the RM2 compared to the
$Re$-independence of the higher $c/R$ UNH-RVAT could be due to this virtual
camber effect, since camber has been shown to cause earlier $Re$-convergence of
the CFT's geometric torque coefficient when calculated from static foil
coefficients given by a viscous panel method, whereas characteristics like
lift-to-drag ratio do not converge as strongly \cite{Bachant2016-RVAT-Re-dep}.

@@

Performance curves for the rotor with both NACA 0021 and cylindrical struts are
shown in Fig.~\ref{fig:perf-covers}. As expected, the high drag cylindrical
struts reduce performance dramatically, producing an entirely negative $C_P$
curve. The detrimental effects of the strut drag were more pronounced at higher
$\lambda$. This is in accordance with the measurements of Rawlings
\cite{Rawlings2008}, though the strut losses in this case were larger due to the
very high drag circular profile. In contrast to the dramatic effect strut drag
had on $C_P$, overall rotor drag measurements were relatively similar for both
the cylindrical and NACA 0021 strut cases.

@@

Measurements for the power coefficient contributions of the strut drag losses
are presented in Fig.~\ref{fig:no-blades} for NACA 0021 and cylindrical
struts---both in towed and stationary conditions. These are computed in the same
fashion as the curves in Fig.~\ref{fig:cp-curves}, but with the rotor blades
removed. We see that strut drag losses increase with tip speed ratio to the
power 2--3, which makes streamlined struts much more important for low solidity
turbines, given the inverse correlation between solidity and $\lambda_0$
\cite{Templin1974}.

@@

Strut drag losses did not change much for the streamlined NACA 0021 struts in
the towed versus stationary configuration, which helps explain why overall rotor
drag coefficients remained of comparable magnitude even at low Reynolds number,
where $C_P$ was dramatically reduced, or even negative. For the cylindrical
struts, losses increased significantly when towed and operating in the mid range
of tip speed ratios.

Though a real turbine would never use such high drag struts, with respect to
numerical modeling, these measurements provide some interesting validation
cases. Modelers can isolate and evaluate the ability to predict these losses
independent of the blade loading by modeling combinations of the rotor with the
high/low drag struts and with/without blades.

@@

The mean velocity field at the chosen optimal tip speed ratio $\lambda_0=3.1$, 1
m downstream ($x/D=0.93$) from the rotor axis is plotted in
Fig.~\ref{fig:meancontquiv}. The mean streamwise velocity deficit is markedly
more symmetric than that of the higher solidity UNH-RVAT, shown in
Fig.~\ref{fig:rvat-meancontquiv}, with the RM2 inducing lower acceleration
around the turbine due to the lower rotor drag coefficient and a slightly lower
blockage ratio \cite{Bachant2015-JoT}. Tip vortex shedding is relatively weaker,
which is likely an effect of the RM2's smaller blade chord length and tapered
blades.

@@

Analysis of the near-wake of the higher solidity UNH-RVAT turbine revealed that
vertical advection was the largest contributor to streamwise recovery
\cite{Bachant2015-JoT}, a trait which is considered an advantage of cross-flow
over axial-flow rotors in arrays \cite{Kinzel2012}. A similar analysis was
undertaken for the RM2, starting with the transport equation for mean kinetic
energy, rearranged to isolate the streamwise partial derivative (indices follow
the Einstein summation convention):
\begin{equation}
    \begin{split}
        \frac{\p K}{\p x}
        =
        \frac{1}{U}
        \bigg{[}
        & - \underbrace{V \frac{\p K}{\p y}}_{y\text{-adv.}}
        - \underbrace{W \frac{\p K}{\p z}}_{z\text{-adv.}}
        % Pressure work:
        - \frac{1}{\rho}\frac{\p}{\p x_j} P U_i \delta_{ij}
        % Work by viscous forces
        + \frac{\p}{\p x_j} 2 \nu U_i S_{ij}
        % Turbulent transport of K
        - \underbrace{
            \frac{1}{2}\frac{\p}{\p x_j} \overline{u_i' u_j'} U_i
        }_{\text{Turb. trans.}} \\
        % Production of k
        & + 
        \underbrace{
            \overline{u_i' u_j'} \frac{\p U_i}{\p x_j}
        }_{k\text{-prod.}}
        % Mean dissipation
        - 
        \underbrace{
            2 \nu S_{ij}S_{ij}
        }_{\text{Mean diss.}}
        \bigg{]}.
    \end{split}
    \label{eq:K-full}
\end{equation}

@@

Derivatives were computed with a second order central difference scheme for
interior points, and a second order inward-facing scheme for the edges,
following the methodology in \cite{Bachant2015-JoT}. Weighted averages for these
calculations are shown in Fig.~\ref{fig:Ktransport}. Due to the weaker blade
vortex shedding, transport due to vertical advection at this point in the wake
was approximately 3 times lower than the higher solidity UNH-RVAT. Note that
direct comparison is somewhat invalid, since the total measurement plane area
was about 5\% lower for this experiment. However, the differences observed are
larger than the area ratio between the two experiments. We also see relatively
lower levels of cross-stream turbulent transport due to the lack of blade stall
vortex shedding. These results may have interesting implications regarding the
application of turbines with lower power coefficient to possibly improve overall
array performance through enhanced transport of kinetic energy from the free
stream, though evaluating these trade-offs will require a detailed analysis of
the downstream evolution and turbine--wake interaction.
